Cross Validation Approaches for Penalized Cox Regression

### 2019-04-02 main.tex
edit 2.3, 3.1, change x label for all figures

The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model. Hence, another approach to cross-validation would be to include an extra step of estimating the baseline hazard and using it to quantify the cross-validation error. The cumulative baseline hazard $\Lambda_0$ is typically estimated via the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
\begin{equation}
  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk at time }t^{-}_j}.
\end{equation}
\note{need to define $t^{-}_j$.}
The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
\begin{equation}
  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\hat\beta).
\end{equation}

@@

A natural candidate for incorporating the cumulative hazard into a loss function is the deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}. First proposed for model diagnostics, deviance residuals have also been occasionally been used as a predictive accuracy measure \citep{Therneau2018}.  Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual: 
\begin{equation}
  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\hat\beta^{-k}).
\end{equation}
The cross-validated deviance residuals can then be derived from the Martingale residuals: 
\begin{equation} 
  d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
\end{equation}
The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$ are used as the cross validated error as an analogy to the residual sums of square for linear regression. We refer to this method as the \emph{cross-validated deviance residuals} approach.

The baseline hazard estimate we describe here differs from the one typically used in constructing deviance residuals \citep{Therneau1990}. Instead of the Nelson-Aalen estimator, deviance residuals are typically calculated based on a baseline hazard estimate that has been adjusted for the covariates in the model such as the one described in \citet{breslow1972}. In the context of cross-validation, however, this approach is problematic. Deviance residuals measure the difference between the fitted model and a saturated model; in Cox regression, this saturated model depends on the baseline hazard. Thus, a covariate-adjusted baseline hazard would mean that each fold is compared to a different saturated model. For this reason, we consider it important that the baseline hazard remains constant across folds when calculating cross-validated deviance residuals; this intuition is borne out by simulations involving various other approaches \note{supplement? otherwise, this sentence is a little tricky}.  Note that this is not the case for the Brier Score or Kullback Leibler Score (described in Section~\ref{Sec:accuracy}), which involve absolute loss functions, not comparisons with a saturated model.

@@

\begin{figure}
  \centering
  \includegraphics[height= 12cm ]{./figures/figure_1_new.png}
  \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
\end{figure}

@@

Throughout, we communicate the estimation accuracy in terms of relative MSE, as compared to the oracle model.  For each generated data set, an oracle model was fit using Cox regression only including the variables with non-zero coefficients. Letting $\hat{\beta}^{oracle}$ denote the resulting estimates, the relative MSE, on the log scale, is given by
\begin{equation}
\log\left\{ \frac{\MSE(\hat{\beta}(\lambda^{\text{cv}}))}{\MSE(\hat{\beta}^{\text{oracle}})} \right\},
\end{equation}
where $\lambda^{cv}$ denotes the value of $\lambda$ selected by cross-validation.   These results are then averaged over N replications; the $\log$ of the ratio is taken to correct for skewness. \bnote{clarification: MSE is estimated by avevaring across N replications respectively for $\hat{\beta}(\lambda^{cv})$ and $\hat{\beta}^{oracle}$. The ratio and $\log$ were taken afterwards } \note{I'm not sure why this is a note; shouldn't you rewrite the sentence?}

%For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$, which we denote by $\lambda^{opt}$. Then the models selected by the cross validation methods with $\lambda^{cv}$ would be compared to the model that corresponds to $\lambda^{opt}$. We would take the ratio between $MSE$ of $\hat{\beta}(\lambda^{cv})$ and $MSE$ of $\hat{\beta}(\lambda^{optimal})$ for each cross-validation selected $\lambda^{cv}$ in each generated data set. The results would be averaged over N replications.

@@

\bnote{I rewrote this paragraph} For each simulated data set, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates. The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11,12,..,1000\}$ and $\beta_j = s$ for $j \in \{1,2, ..., 10 \}$, where $s$ is a prespecified constant that parameterizes the signal strength in the simulated data set. In the first simulation study, we varied $s$ from 0.4 to 0.9. The survival outcomes have 10$\%$ expected censoring. For each of the four methods described in Section~\ref{Sec:methods}, 10-fold cross-validation was implemented. $N = 200$ replications were used for each scenario.

>>>>

The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model. Hence, another approach to cross-validation would be to include an extra step of estimating the baseline hazard and using it to quantify the cross-validation error. The cumulative baseline hazard $\Lambda_0$ is typically estimated via the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
\begin{equation}
  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
\end{equation}
\bnote{I write out the $t^{-}_j$ definition directly into the fomula so that we don't need to define $t^{-}_j$ separately.}
The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
\begin{equation}
  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\hat\beta).
\end{equation}

@@

A natural candidate for incorporating the cumulative hazard into a loss function is the deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}. First proposed for model diagnostics, deviance residuals have also been occasionally used as a predictive accuracy measure \citep{Therneau2018}.  Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual: 
\begin{equation}
  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\hat\beta^{-k}).
\end{equation}
The cross-validated deviance residuals can then be derived from the Martingale residuals: 
\begin{equation} 
  d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
\end{equation}
The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$ are used as the cross validated error as an analogy to the residual sums of square for linear regression. We refer to this method as the \emph{cross-validated deviance residuals} approach.

The baseline hazard estimate we describe here differs from the one typically used in constructing deviance residuals \citep{Therneau1990}. Instead of the Nelson-Aalen estimator, deviance residuals are typically calculated based on a baseline hazard estimate that has been adjusted for the covariates in the model such as the one described in \citet{breslow1972}. In the context of cross-validation, however, this approach is problematic. Deviance residuals measure the difference between the fitted model and a saturated model; in Cox regression, this saturated model depends on the baseline hazard. Thus, a covariate-adjusted baseline hazard would mean that each fold is compared to a different saturated model. For this reason, we consider it important that the baseline hazard remains constant across folds when calculating cross-validated deviance residuals; this intuition is borne out by simulations involving various other approaches \note{supplement? otherwise, this sentence is a little tricky}\bnote{sure}.  Note that this is not the case for the Brier Score or Kullback Leibler Score (described in Section~\ref{Sec:accuracy}), which involve absolute loss functions, not comparisons with a saturated model.

@@

\begin{figure}
  \centering
  \includegraphics[height= 12cm ]{./manuscript_figure/figure_1_new.png}
  \caption{Illustration of K-fold Cross Validation Methods for Cox Regression}
\end{figure}

@@

Throughout, we communicate the estimation accuracy in terms of relative MSE, as compared to the oracle model.  For each generated data set, an oracle model was fit using Cox regression only including the variables with non-zero coefficients. Letting $\hat{\beta}^{oracle}$ denote the resulting estimates, the relative MSE, on the log scale, is given by
\begin{equation}
\log\left\{ \frac{\MSE(\hat{\beta}(\lambda^{\text{cv}}))}{\MSE(\hat{\beta}^{\text{oracle}})} \right\},
\end{equation}
where $\lambda^{cv}$ denotes the value of $\lambda$ selected by cross-validation. MSE is estimated by avevaring over N replications respectively for $\hat{\beta}(\lambda^{cv})$ and $\hat{\beta}^{oracle}$. The ratio is taken afterwards. The $\log$ is taken to correct for skewness. %\note{I'm not sure why this is a note; shouldn't you rewrite the sentence?}

@@

 For each simulated data set, the covariate matrix has $n = 120$ observations and $p = 1000$ covariates. The covariate vector $\beta$ is assumed to be sparse, with $\beta_j = 0$ for $j \in \{11,12,..,1000\}$ and $\beta_j = s$ for $j \in \{1,2, ..., 10 \}$, where $s$ is a prespecified constant that parameterizes the signal strength in the simulated data set. In the first simulation study, we varied $s$ from 0.4 to 0.9. The survival outcomes have 10$\%$ expected censoring. For each of the four methods described in Section~\ref{Sec:methods}, 10-fold cross-validation was implemented. $N = 200$ replications were used for each scenario.
